MASCARA Playground

### 2023-12-11 R/run_first/1.pattern_examples.Rmd
231211

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(devtools)
library(ggplot2)
library(MetStaT)
library(gASCA)
library(patchwork)
library(readxl)
library(ggfortify)
library(ggrepel)
library(grid)
library(gridExtra)
library(tidyr)
library(tidyverse)
library(reshape2)
library(scales)
library(matrixStats)
library(DESeq2)
library(MASS)
library(glmnet)
library(doParallel)    
library(MUVR) 
library(pheatmap)
library(caret)
library(data.table)
library(plotly)
library(mixOmics)
library(DiscriMiner)
library(pracma)
# library(NineteenEightyR)
library(matlib)
library(truncnorm)
# library(sf)

source("DATA_SIM_FUNCS.R")

@@

set.seed(3)
rnd_seed <- abs(rnorm(3, mean = 0.7, sd = 0.5))
#c(1,0.8,0.3)
# Create_Core_DEV
sim_data <- Create_Core_DEV_2(nreps = 3, meta = meta, plot = TRUE, 
                        EffectSize = c(X_a_ab = 5, time = 1, E = 10, mu = 1, Struc_E = 1), 
                        struc_resid = TRUE, e_sigma = c(1.5,0.8,0.0), a_sigma = c(1,0.7,0.6,0.1),
                        b_sigma = c(1,0.9), irr_spikes = TRUE, SCORE_SEED = 12, Experiment_responders = Experiment_responders,
                        struc_seed = 127)  #12

#list(X,P,Pb,a_ab,PLOT, X_a_ab,time, Struc_E )

###plot error in sim_data[[5]]

@@

# sim_data[[1]] <- sim_data[[1]][,1850:2000]

# image(sim_data)
```

@@

# data <- (0.1 * interaction) + (1.5*time) + (0.5*ambient) + (0*noise)
# 
input <- data
# 
# input <- cbind.data.frame(input,input,input)

@@

```{r}
# data <- (0.1 * interaction) + (1.5*time) + ambient + noise

# input <- input + rnorm(nrow(input) * ncol(input),mean = 0, sd = 1)
#replace with simulation input
# colnames(input) <- make.names(colnames(input))

@@

```{r}
MASCARA4_test_1 <- function(resids,ref, baits, spikes, ncands){

  Fref <- ref

  # spls_res <- simpls.fit(resids[,-which(colnames(resids) %in% baits)], 
  #                  resids[,which(colnames(resids) %in% baits)], all.outputs = TRUE)

  sPLS_cands <- cbind.data.frame(abs(spls_res$loadings$X) %*% spls_res$prop_expl_var$X, spls_res$loadings$X) #  * spls_res$prop_expl_var$X
  sPLS_cands <- sPLS_cands[order(sPLS_cands[,1], decreasing = TRUE),]

  importance <- paste0(round(spls_res$prop_expl_var$X * 100, digits = 2), "%")

  Fref2 <- Fref[-which(Fref$Feature %in% baits),]
  Fref2 <- Fref2[match(rownames(sPLS_cands),Fref2$Feature),]

  #####

  y <- colMeans(spls_res$loadings$Y)

  U <- spls_res$loadings$X

  # UTP <- ((U %*% y)/norm(y, "2")) * y
  # UTP <- ((U %*% y)/dot(y,y)) * y   #_9

  # UTP <- (dot(t(U),y)/dot(y,y)) * y  #_10

  UTP <- (dot(t(U),y)/dot(y,y))

  sPLS_bait_cands <- as.data.frame(UTP[order(UTP, decreasing = TRUE)])

  # sPLS_bait_cands <- as.data.frame(UTP[order(UTP[,1], decreasing = TRUE),])  #_9 and below

  sPLS_maxima_cands <- sPLS_bait_cands

  Fref2 <- Fref[-which(Fref$Feature %in% baits),]
  Fref2 <- Fref2[match(rownames(sPLS_maxima_cands),Fref2$Feature),]

  RP <- prod(which(rownames(sPLS_maxima_cands) %in% spikes))^(1/length(spikes))

  return(list(RP, sPLS_maxima_cands,UTP))
}

MASCARA4_test_2 <- function(resids,ref, baits, spikes, ncands){

  Fref <- ref

  # spls_res <- simpls.fit(resids[,-which(colnames(resids) %in% baits)], 
  #                  resids[,which(colnames(resids) %in% baits)], all.outputs = TRUE)

  sPLS_cands <- cbind.data.frame(abs(spls_res$loadings$X) %*% spls_res$prop_expl_var$Y, spls_res$loadings$X) #  * spls_res$prop_expl_var$X
  sPLS_cands <- sPLS_cands[order(sPLS_cands[,1], decreasing = TRUE),]

  importance <- paste0(round(spls_res$prop_expl_var$X * 100, digits = 2), "%")

  Fref2 <- Fref[-which(Fref$Feature %in% baits),]
  Fref2 <- Fref2[match(rownames(sPLS_cands),Fref2$Feature),]

  #####

  y <- colMeans(spls_res$loadings$Y)

  U <- spls_res$loadings$X

  # UTP <- ((U %*% y)/norm(y, "2")) * y
  # UTP <- ((U %*% y)/dot(y,y)) * y   #_9

  # UTP <- (dot(t(U),y)/dot(y,y)) * y  #_10

  UTP <- (dot(t(U),y)/dot(y,y))

  sPLS_bait_cands <- as.data.frame(UTP[order(UTP, decreasing = TRUE)])

  # sPLS_bait_cands <- as.data.frame(UTP[order(UTP[,1], decreasing = TRUE),])  #_9 and below

  sPLS_maxima_cands <- sPLS_bait_cands

  Fref2 <- Fref[-which(Fref$Feature %in% baits),]
  Fref2 <- Fref2[match(rownames(sPLS_maxima_cands),Fref2$Feature),]

  RP <- prod(which(rownames(sPLS_maxima_cands) %in% spikes))^(1/length(spikes))

  return(list(RP, sPLS_maxima_cands,UTP))
}

```{r}
##feature selection alternatives
source("MASCARA_Test_Funcs.R")
#default - target projection onto average Q in 2 dimensions
resids <- ar[[3]]

@@

#PLS1 model with svd(baits)$T as response
MASCARA_3 <- MASCARA4_test_3(ar[[3]], baits, spikes = spikes)
which(rownames(MASCARA_3[[2]]) %in% spikes)

# 
# 
#

#PLS1 model with svd(baits)$T as response and target projected deflation

# MASCARA_4 <- MASCARA4_test_4(ar[[3]], baits, spikes = spikes)
# which(rownames(MASCARA_4[[2]]) %in% spikes)

# #PLS2 model with target projected deflation
# 
# 
#

@@

```{r}
source("MASCARA_Test_Funcs.R")
#classic
MASCARA_2 <- MASCARA4_test_2(ar[[3]], baits, spikes = spikes)
which(rownames(MASCARA_2[[2]]) %in% spikes)

#u1 of baits as pls1 response
MASCARA_3 <- MASCARA4_test_3(ar[[3]], baits, spikes = spikes)
which(rownames(MASCARA_3[[2]]) %in% spikes)

@@

#vip

MASCARA_5 <- MASCARA4_test_8(ar[[3]], baits, spikes = spikes)
which(rownames(MASCARA_5[[2]]) %in% spikes)

#tpsr

@@

# yloadings[3,2] <- -0.5*yloadings[3,2]   #pure data manipulation for demo purposes

@@

# loadings_r_baits <- rbind(loadings_r_baits, yloadings)

@@

    scale_color_manual(values=c('#E69F00',"black", '#56B4E9'))+
      geom_point(data = loadings_r_baits[which(loadings_r_baits$ID %in% c("Baits","Pathway")),], aes(size = 15)) +
  # geom_point(data = yloadings, shape = 4, aes(x = Comp_1, y = Comp_2, size = 3 )) + 
  theme_bw()

@@

#multiple PLS1 models with svd(baits) deflated

```{r}
#remake simpls to deflate with target projected loadings

#first recreate simpls such that the output is the same as simpls.fit()

#$projection -> T = XR

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = "../")
```

```{r, echo=FALSE}
library(devtools)
library(ggplot2)
library(MetStaT)
library(gASCA)
library(patchwork)
library(readxl)
library(ggfortify)
library(ggrepel)
library(grid)
library(gridExtra)
library(tidyr)
library(tidyverse)
library(reshape2)
library(scales)
library(matrixStats)
library(DESeq2)
library(MASS)
library(glmnet)
library(doParallel)    
library(MUVR) 
library(pheatmap)
library(caret)
library(data.table)
library(plotly)
library(mixOmics)
library(DiscriMiner)
library(pracma)
library(matlib)
library(truncnorm)

source("../DATA_SIM_FUNCS.R")
```

@@

set.seed(3)
rnd_seed <- abs(rnorm(3, mean = 0.7, sd = 0.5))

sim_data <- Create_Core_DEV_2(nreps = 3, meta = meta, plot = TRUE, 
                        EffectSize = c(X_a_ab = 5, time = 1, E = 10, mu = 1, Struc_E = 1), 
                        struc_resid = TRUE, e_sigma = c(1.5,0.8,0.0), a_sigma = c(1,0.7,0.6,0.1),
                        b_sigma = c(1,0.9), irr_spikes = TRUE, SCORE_SEED = 12, Experiment_responders = Experiment_responders,
                        struc_seed = 127)  
```

@@

@@

input <- data

@@

@@

```{r}
##feature selection alternatives
source("../MASCARA_FUNCS.R")
#default - target projection onto average Q in 2 dimensions
resids <- ar[[3]]

@@

@@

```{r}
source("../MASCARA_FUNCS.R")
#classic
MASCARA_2 <- MASCARA4_test_2(ar[[3]], baits, spikes = spikes)
which(rownames(MASCARA_2[[2]]) %in% spikes)

@@

#vip
MASCARA_5 <- MASCARA4_test_8(ar[[3]], baits, spikes = spikes)
which(rownames(MASCARA_5[[2]]) %in% spikes)

@@

@@

@@

    scale_color_manual(values=c('#E69F00',"black", '#56B4E9'))+
      geom_point(data = loadings_r_baits[which(loadings_r_baits$ID %in% c("Baits","Pathway")),], aes(size = 15)) +
  theme_bw()

@@

### 2023-12-11 R/run_first/1.pattern_examples_MASCARA_pipeline.Rmd
231211

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = "../")
```

@@

source("../DATA_SIM_FUNCS.R")
```

@@

set.seed(3)
rnd_seed <- abs(rnorm(3, mean = 0.7, sd = 0.5))

sim_data <- Create_Core_DEV_2(nreps = 3, meta = meta, plot = TRUE, 
                        EffectSize = c(X_a_ab = 5, time = 1, E = 10, mu = 1, Struc_E = 1), 
                        struc_resid = TRUE, e_sigma = c(1.5,0.8,0.0), a_sigma = c(1,0.7,0.6,0.1),
                        b_sigma = c(1,0.9), irr_spikes = TRUE, SCORE_SEED = 12, Experiment_responders = Experiment_responders,
                        struc_seed = 127)  
```

@@

sim_data <- Create_Core_DEV_2(nreps = 3, meta = meta, 
                        EffectSize = c(X_a_ab = 5, time = 1, E = 1, mu = 1, Struc_E = 0.4),    # E = 10, Struc_E = 1
                        struc_resid = TRUE, e_sigma = c(1.5,0.8,0.0), a_sigma = c(1,0.7,0.6,0.1),
                        b_sigma = c(1,0.9), irr_spikes = TRUE, SCORE_SEED = 12, Experiment_responders = Experiment_responders,
                        struc_seed = 127, score_plot = FALSE, plot = FALSE)  #12

@@

ref <- cbind.data.frame(Feature = colnames(sim_data[[1]]), sim_data[[2]], sim_data[[3]])
cols <- colnames(ref[,-1])
ref$Effect <- do.call(paste, c(ref[cols], sep = "_"))
ref$Effect[which(ref$Effect == "0.5_0_0_0.1")] <- "Low PC1"
ref$Effect[which(ref$Effect == "1_0_0_0.1")] <- "Medium PC1"
ref$Effect[which(ref$Effect == "1.5_0_0_0.1")] <- "High PC1"
ref$Effect[which(ref$Effect == "0_0.5_0.1_0")] <- "PC2"
ref$Effect[grep("^0", ref$Effect)] <- "None"
ref$Effect <- factor(ref$Effect, levels = c("Low PC1", "Medium PC1", "High PC1", "PC2", "None"))
colnames(ref)[8] <- "Baits"

ref$Description <- ref$Baits

ref$Baits[1985:1988] <- "Baits"
ref$Baits[1989:2000] <- "Candidates"

ref$Baits[(2000 - (16 + Experiment_responders)):1984] <- "ab1+ve"
# ref$Baits[1969:1976] <- "ab1+ve"
ref$Baits[501:515] <- "ab2+ve"
ref$Baits[486:500] <- "ab2-ve"
# ref$Baits[1969:1976] <- "Expt_Independent"

ref$Baits[1:24] <- "ab1-ve"

ref$Description <- ref$Baits

@@

library(viridis)
library(hrbrthemes)
library(ggplot2)
p1 <- ggplot(mData, aes(Samples, Features, fill= value)) + 
  geom_tile() +
  ggtitle("Growth Condition") +
  scale_fill_viridis(discrete=FALSE) +  #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

mData <- reshape2::melt(time)
colnames(mData)[1:2] <- c("Samples","Features")
# mData$Features <- fct_rev(mData$Features)

library(viridis)
library(hrbrthemes)
library(ggplot2)
p2 <- ggplot(mData, aes(Samples, Features, fill= value)) + 
  geom_tile() +
  ggtitle('Time') +
  scale_fill_viridis(discrete=FALSE) +  #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

@@

mData <- reshape2::melt(interaction)
colnames(mData)[1:2] <- c("Samples","Features")
# mData$Features <- fct_rev(mData$Features)

library(viridis)
library(hrbrthemes)
library(ggplot2)
p3 <- ggplot(mData, aes(Samples, Features, fill= value)) + 
  geom_tile() +
  ggtitle('Interaction') +
  scale_fill_viridis(discrete=FALSE) +  #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

mData <- reshape2::melt(ambient)
colnames(mData)[1:2] <- c("Samples","Features")
# mData$Features <- fct_rev(mData$Features)

library(viridis)
library(hrbrthemes)
library(ggplot2)
p4 <- ggplot(mData, aes(Samples, Features, fill= value)) + 
  geom_tile() +
  ggtitle('Ambient') +
  scale_fill_viridis(discrete=FALSE ) +   #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

mData <- reshape2::melt(noise)
colnames(mData)[1:2] <- c("Samples","Features")
# mData$Features <- fct_rev(mData$Features)

library(viridis)
library(hrbrthemes)
library(ggplot2)
p5 <- ggplot(mData, aes(Samples, Features, fill= value)) + 
  geom_tile() +
  ggtitle('Residuals') +
  scale_fill_viridis(discrete=FALSE) +  #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

data <- (0.1 * interaction) + (1.5*time) + ambient + noise
mData <- reshape2::melt(data)
colnames(mData)[1:2] <- c("Samples","Genes")
# mData$Features <- fct_rev(mData$Features)

library(viridis)
library(hrbrthemes)
library(ggplot2)
p0 <- ggplot(mData, aes(Samples, Genes, fill= value)) + 
  geom_tile() +
  ggtitle('Data') +
  scale_fill_viridis(discrete=FALSE) +  #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank())#,
        #axis.title.x = element_blank(),
        #axis.title.y = element_blank())

@@

```{r}
ps <- patchwork::patchworkGrob(wrap_plots(plots) + plot_layout(guides = "collect"))  #, heights = unit(c(5), c("cm")), 
ps1 <- grid.arrange(ps)#, left = "Features", bottom = "Replicate Number                     ")

# p5
```

@@

library(viridis)
library(hrbrthemes)
library(ggplot2)
p5 <- ggplot(mData, aes(Samples, Features, fill= value)) + 
  geom_tile() +
  ggtitle('Residuals') +
  scale_fill_viridis(discrete=FALSE) +  #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

@@

input <- data

@@

```{r}
####PLS pattern visualisation change this

data <- scores[]    # data.matrix(sim_data[[4]][,3:4])  #
mData <- reshape2::melt(data)
colnames(mData)[1:2] <- c("Samples","Features")
# mData$Features <- fct_rev(mData$Features)

library(viridis)
library(hrbrthemes)
library(ggplot2)
p0 <- ggplot(mData, aes(Features,Samples, fill= value)) + 
  geom_tile() +
  ggtitle('Patterns (T)') +
  scale_fill_viridis(discrete=FALSE) +  #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none")

# p0

data <- input[,1:2]
mData <- reshape2::melt(data)
colnames(mData)[1:2] <- c("Samples","Features")
# mData$Features <- fct_rev(mData$Features)

library(viridis)
library(hrbrthemes)
library(ggplot2)
p1 <- ggplot(mData, aes(Samples,Features, fill= value)) + 
  geom_tile() +
  ggtitle('Data; X') +
  scale_fill_viridis(discrete=FALSE) +  #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none")

data <- input[,3:4]
mData <- reshape2::melt(data)
colnames(mData)[1:2] <- c("Samples","Features")
# mData$Features <- fct_rev(mData$Features)

library(viridis)
library(hrbrthemes)
library(ggplot2)
p2 <- ggplot(mData, aes(Samples,Features, fill= value)) + 
  geom_tile() +
  ggtitle('Data; Y') +
  scale_fill_viridis(discrete=FALSE) +  #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none")

p0 + (p1 / p2) + theme(legend.position = "none")
```

```{r}
##feature selection alternatives
source("../MASCARA_FUNCS.R")
#default - target projection onto average Q in 2 dimensions
resids <- ar[[3]]

MASCARA_1 <- MASCARA4_test_1(ar[[3]], baits = baits, spikes = spikes)
which(rownames(MASCARA_1[[2]]) %in% spikes)

# # replacing mixomics with SIMPLS from pls package

#1 component, absolute projections
spls_res2.1 <- simpls.fit(resids[,-which(colnames(resids) %in% baits)],
                   resids[,which(colnames(resids) %in% baits)], ncomp = 1)

res_prj <- cbind.data.frame(rownames(spls_res2.1$projection),spls_res2.1$projection)
res_VIP <- res_prj[order(abs(res_prj[,2]), decreasing = TRUE),]
rownames(res_VIP) <- res_VIP[,1]

which(rownames(res_VIP) %in% spikes)

## 2 component target projection
MASCARA_2 <- MASCARA4_test_2(ar[[3]], baits, spikes = spikes)
which(rownames(MASCARA_2[[2]]) %in% spikes)

```{r}
resids <- ar[[3]]

spls_res <- spls(resids[,-which(colnames(resids) %in% baits)], 
                 resids[,which(colnames(resids) %in% baits)], all.outputs = TRUE)

spls_res2 <- simpls.fit(resids[,-which(colnames(resids) %in% baits)],
                resids[,which(colnames(resids) %in% baits)], ncomp = 2)

spls_res2.1 <- simpls.fit(resids[,-which(colnames(resids) %in% baits)],
                   resids[,which(colnames(resids) %in% baits)], ncomp = 1)

res_prj <- cbind.data.frame(rownames(spls_res2.1$projection),spls_res2.1$projection)
res_VIP <- res_prj[order(abs(res_prj[,2]), decreasing = TRUE),]
rownames(res_VIP) <- res_VIP[,1]

which(rownames(res_VIP) %in% spikes)

```{r}
source("../MASCARA_FUNCS.R")
#classic
MASCARA_2 <- MASCARA4_test_2(ar[[3]], baits, spikes = spikes)
which(rownames(MASCARA_2[[2]]) %in% spikes)

@@

# str(MASCARA_2[[4]])

@@

@@

LINE <- rbind(line, c(0,0))
LINE <- t(LINE)
colnames(LINE) <- c("to","from")
# scores <- scores %*% 
scores <- cbind.data.frame(scores,meta[,1:2])

@@

loadings_r_baits$ID[which(loadings_r_baits$ID %in% spikes)] <- "Pathway"
loadings_r_baits$ID[which(loadings_r_baits$ID %in% baits)] <- "Baits"
```

@@

    scale_color_manual(values=c('#E69F00',"black", '#56B4E9'))+
      geom_point(data = loadings_r_baits[which(loadings_r_baits$ID %in% c("Baits","Pathway")),], aes(size = 15)) +
  theme_bw()

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

source("../DATA_SIM_FUNCS.R")

@@

@@

sim_data <- Create_Core_DEV_2(nreps = 3, meta = meta, 
                        EffectSize = c(X_a_ab = 5, time = 1, E = 10, mu = 1, Struc_E = 1),    # E = 1, Struc_E = 0.4
                        struc_resid = TRUE, e_sigma = c(1.5,0.8,0.0), a_sigma = c(1,0.7,0.6,0.1),
                        b_sigma = c(1,0.9), irr_spikes = TRUE, SCORE_SEED = 12, Experiment_responders = Experiment_responders,
                        struc_seed = 127, score_plot = FALSE, plot = FALSE)

@@

@@

library(viridis)
library(hrbrthemes)
library(ggplot2)
p1 <- ggplot(mData, aes(Features, Samples, fill= value), show.legend = FALSE) + 
  geom_tile() +
  ggtitle("Growth Condition") +
  scale_fill_viridis(discrete=FALSE) +  
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

mData <- reshape2::melt(time)
colnames(mData)[1:2] <- c("Samples","Features")

library(viridis)
library(hrbrthemes)
library(ggplot2)
p2 <- ggplot(mData, aes(Features, Samples, fill= value)) + 
  geom_tile() +
  ggtitle('Time') +
  scale_fill_viridis(discrete=FALSE) +  #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

@@

mData <- reshape2::melt(interaction)
colnames(mData)[1:2] <- c("Samples","Features")

library(viridis)
library(hrbrthemes)
library(ggplot2)
p3 <- ggplot(mData, aes(Features, Samples, fill= value)) + 
  geom_tile() +
  ggtitle('Interaction') +
  scale_fill_viridis(discrete=FALSE) +  #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

mData <- reshape2::melt(ambient)
colnames(mData)[1:2] <- c("Samples","Features")

library(viridis)
library(hrbrthemes)
library(ggplot2)
p4 <- ggplot(mData, aes(Features, Samples, fill= value)) + 
  geom_tile() +
  ggtitle('Ambient') +
  scale_fill_viridis(discrete=FALSE ) +   #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

mData <- reshape2::melt(noise)
colnames(mData)[1:2] <- c("Samples","Features")

library(viridis)
library(hrbrthemes)
library(ggplot2)
p5 <- ggplot(mData, aes(Features, Samples, fill= value)) + 
  geom_tile() +
  ggtitle('Residuals') +
  scale_fill_viridis(discrete=FALSE) +  #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

data <- (0.1 * interaction) + (1.5*time) + ambient + noise
mData <- reshape2::melt(data)
colnames(mData)[1:2] <- c("Samples","Genes")

library(viridis)
library(hrbrthemes)
library(ggplot2)
p0 <- ggplot(mData, aes(Genes, Samples, fill= value)) + 
  geom_tile() +
  ggtitle('Data') +
  scale_fill_viridis(discrete=FALSE) +  #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16))

@@

p5

```{r}
ps <- patchwork::patchworkGrob(wrap_plots(plots) )    #+ plot_layout(guides = "collect") -

ps1 <- grid.arrange(ps)

@@

library(viridis)
library(hrbrthemes)
library(ggplot2)
p5 <- ggplot(mData, aes(Features, Samples, fill= value)) + 
  geom_tile() +
  ggtitle('Residuals') +
  scale_fill_viridis(discrete=FALSE) +  #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

@@

# 
input <- data
#

@@

```{r}
source("../MASCARA_FUNCS.R")

#classic; q bar target projection
MASCARA_2 <- MASCARA4_test_2(ar[[3]], baits, spikes = spikes)
which(rownames(MASCARA_2[[2]]) %in% spikes)

#u1 of baits as pls1 response
MASCARA_3 <- MASCARA4_test_3(ar[[3]], baits, spikes = spikes)
which(rownames(MASCARA_3[[2]]) %in% spikes)

@@

@@

#rotation of scores and loadings for demo purposes only

@@

LINE <- rbind(line, c(0,0))
LINE <- t(LINE)
colnames(LINE) <- c("to","from")

scores <- cbind.data.frame(scores,meta[,1:2])

@@

loadings_r_baits$ID[which(loadings_r_baits$ID %in% spikes)] <- "Spikes"
loadings_r_baits$ID[which(loadings_r_baits$ID %in% baits)] <- "Baits"
```

@@

    scale_color_manual(values=c('#E69F00',"black", '#56B4E9'))+
      geom_point(data = loadings_r_baits[which(loadings_r_baits$ID %in% c("Baits","Spikes")),], aes(size = 15)) +
  theme_bw()
